### 2022-07-20 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@2938f44d869e03e1f18265e29ad281d36f52bd1b 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=4.5cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\thispagestyle{empty}
{\Large\bfseries\sffamily What constrains food webs? A maximum entropy
framework for predicting their structure with minimal biases}
\vskip 5em

%
\href{https://orcid.org/0000-0001-9051-0597}{Francis\,Banville}%
%
\,\textsuperscript{1,2,3}\quad %
\href{https://orcid.org/0000-0002-4498-7076}{Dominique\,Gravel}%
%
\,\textsuperscript{2,3}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,3}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Université de
Sherbrooke\quad \textsuperscript{3}\,Quebec Centre for Biodiversity
Science

\textbf{Correspondance to:}\\
Francis Banville --- \texttt{francis.banville@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
\textbf{\sffamily Abstract: }Food webs are complex ecological networks
whose structure is both ecologically and statistically constrained, with
many network properties being correlated with each other. Despite the
recognition of these invariable relationships in food webs, the use of
the principle of maximum entropy (MaxEnt) in network ecology is still
rare. This is surprising considering that MaxEnt is a renowned and
rigorous statistical tool precisely designed for understanding and
predicting many different types of constrained systems. Precisely, this
principle asserts that the least-biased probability distribution of a
system's property, constrained by prior knowledge about that system, is
the one with maximum information entropy. MaxEnt has been proven useful
in many ecological modelling problems, but its application in food webs
and other ecological networks is limited. Here we show how MaxEnt can be
used to derive many food-web properties both analytically and
heuristically. First, we show how the joint degree distribution (the
joint probability distribution of the numbers of prey and predators for
each species in the network) can be derived analytically using the
number of species and the number of interactions in food webs. Second,
we present a heuristic and flexible approach of finding a network's
adjacency matrix (the network's representation in matrix format) based
on simulated annealing and SVD entropy. We built two heuristic models
using the connectance and the joint degree sequence as statistical
constraints, respectively. We compared both models' predictions against
corresponding null and neutral models commonly used in network ecology
using open access data of terrestrial and aquatic food webs sampled
globally (N = 257). We found that the heuristic model constrained by the
joint degree sequence was a good predictor of many measures of food-web
structure, especially the nestedness and motifs distribution.
Specifically, our results suggest that the structure of terrestrial and
aquatic food webs is mainly driven by their joint degree distribution.
\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Statistical and mathematical models can help fill many gaps in our
knowledge about species interactions. Predictive and null models are two
complementary types of models that have been developed in network
ecology for this purpose. On one hand, predictive models can partially
alleviate the Eltonian shortfall, which describes our current lack of
knowledge on food webs and other ecological networks (Hortal et al.
2015). A variety of such models have recently been developed using
machine learning and other statistical tools, most of which are
presented in Strydom et al. (2021). On the other hand, null models help
us identify potential ecological mechanisms that drive species
interactions. They do so by comparing empirical data with an unbiased
distribution of measures generated using a set of rules that exclude the
mechanism of interest (Fortuna and Bascompte 2006; Delmas et al. 2019).
Both types of models are frequently topological, i.e.~they often predict
the adjacency matrix or specific measures of network structure without
taking into account species' identity. According to Strydom et al.
(2021), these topological models could be used to make better
predictions of pairwise species interactions by constraining the space
of feasible networks.

The principle of maximum entropy (MaxEnt) is a statistical and
topological model that can be used for both of these purposes, i.e.~to
make predictions of network structure and to better understand processes
shaping ecological networks. Food-web properties that can be derived
using MaxEnt are varied and pertain to different elements of the network
(i.e.~at the species (node), the interaction (edge) or the community
(network) levels). Because MaxEnt is a method of finding probability
distributions that are least biased given partial knowledge about a
system, these properties need to be represented probabilistically.
Otherwise, some data transformation might be needed. For example, at the
species level, MaxEnt can be used to predict the distribution of trophic
levels among species, as well as the distribution of species'
vulnerability (number of predators) and generality (number of preys). By
contrast, at the interaction level, predictions can be made on the
distribution of interaction strengths in weighted food webs. At the
community level, it can give us a probability distribution of potential
networks or of specific measures of their emerging structure. The
applicability and performance of this method depend on the ecological
information available and on our capacity to find the set of state
variables that best represent natural systems and to translate them into
appropriate statistical constraints. MaxEnt has been used in a wide
range of disciplines, from thermodynamics to chemistry and biology
(Martyushev and Seleznev 2006). It has also been proven useful in
ecology, e.g.~in species distribution models (Phillips, Anderson, and
Schapire 2006) and macroecological models (Harte et al. 2008; Harte and
Newman 2014). The probability distribution of the variable of interest
with the maximum entropy value is unique and least biased under the set
of constraints used. These constraints are built using state variables,
i.e.~variables that represent the macrostate of the system. Having a
validated maximum entropy model for the system at hand allows us to make
rigorous predictions using a minimal amount of data, as well as helping
us describe the most important factors driving that system.

Despite its extensive use in graph and network theories (e.g., Park and
Newman 2004; van der Hoorn, Lippner, and Krioukov 2018), MaxEnt has in
comparison been little used in network ecology. Like other real system
networks, food webs and other ecological networks are represented
mathematically as graphs. However, the very nature of food webs
(directed simple graphs frequently having self-loops) makes the
mathematical optimization of maximum entropy graph models more
complicated than with many other types of (non-ecological) networks.
MaxEnt has nevertheless been used to predict the degree distribution of
bipartite ecological networks from the number of species and the number
of interactions (Williams 2011) and to predict interaction strengths
between species pairs using their relative abundances within an optimal
transportation theory regularized with information entropy (Stock,
Poisot, and De Baets 2021). However, to the best of our knowledge,
MaxEnt has never been used to predict food-web structure directly, even
though food webs are among the most documented and widespread ecological
networks.

In this contribution, we used two complementary approaches to predict
the structure of food webs using the principle of maximum entropy. We
then compared our predictions against empirical data and null and
neutral models commonly used in network ecology. The first approach
consists in deriving constrained probability distributions of given
network properties directly. We derived the joint degree distribution
(the joint probability distribution that a species has a given number of
prey and predators in a food web) of maximum entropy using only the
number of species \(S\) and the number of interactions \(L\) as state
variables. Then, we predicted the degree distribution of maximum entropy
directly from the joint degree distribution since the first is the sum
of the marginal distributions of the second (a species' degree is the
sum of its in and out-degrees). Because of the scarcity of empirical
data on the number of interactions in ecological networks, we present a
method to predict \(L\) from \(S\) (Box 1), thus allowing the prediction
of the joint degree distribution from \(S\) solely. In turn, the second
approach consists in finding, using different constraints, the adjacency
matrix of maximum entropy from which network properties can be measured.
To do so, we used a flexible and heuristic approach based on simulated
annealing (an optimization algorithm) to find networks \emph{close} to
maximum entropy. As discussed above, our choice of algorithm stands from
the very nature of food webs (i.e.~simple directed networks allowing
self-loops) that makes the analytical derivation of a maximum entropy
graph model difficult. We first built our type I MaxEnt network model
constrained by the connectance of the network (i.e.~the ratio
\(L/S^2\)). A comparison of this model against empirical data indicated
that connectance alone was not sufficient to predict many aspects of
network structure. For this reason, we built our type II MaxEnt network
model, which instead uses the whole joint degree sequence as a
constraint. Overall, we found that this second model was much better at
predicting food-web structure than the one based on connectance.

\hypertarget{the-principle-of-maximum-entropy-a-primer-for-ecologists}{%
\subsection{The principle of maximum entropy: A primer for
ecologists}\label{the-principle-of-maximum-entropy-a-primer-for-ecologists}}

The principle of maximum entropy is a mathematical method of finding
probability distributions, strongly rooted in statistical mechanics and
information theory (Jaynes 1957a, 1957b; Harremoës and Topsøe 2001).
Starting from a set of constraints given by prior knowledge of a system
(i.e.~what we call state variables), this method helps us find
least-biased probability distributions subject to the constraints. These
probability distributions are guaranteed to be unique given our prior
knowledge and represent the most we can say about a system without
making more assumptions. For example, if the only thing we know about a
biological community is its average number of individuals per species,
the least-biased inference we could make on its species abundance
distribution is the exponential distribution (Frank and Smith 2011;
Harte and Newman 2014). However, this does not imply that this
distribution will be the best fit to empirical data. The challenge is to
find the right set of constraints that would best reproduce
distributions found in nature.

The least-biased probability distribution given the constraints used is
the one with the highest entropy among all probability distributions
that satisfy these constraints. Entropy is a measure of the average
amount of information given by the outcome of a random variable. Many
measures of entropy have been developed in physics (Beck 2009), but only
a fraction of them could be used as an optimization measure with the
principle of maximum entropy. According to Beck (2009) and Khinchin
(2013), a measure of entropy \(H\) should satisfy four properties in the
discrete case: (1) it should be a function of a probability distribution
\(p(n)\) only; (2) it should be maximized when \(p(n)\) is uniform; (3)
it should not be influenced by outcomes with a null probability; and (4)
it should be independent of the order of information acquisition. The
Shannon's entropy (Shannon 1948)

\begin{equation}\protect\hypertarget{eq:shannon}{}{H = -\sum_{n} p(n) \log p(n)}\label{eq:shannon}\end{equation}

satisfies all of these properties. Finding the probability distribution
\(p(n)\) that maximizes \(H\) under a set of \(m\) constraints \(g\) can
be done using the method of Lagrange multipliers. These constraints
could include one or many properties of the probability distribution
(e.g., its mean, variance, and range). However, the normalization
constraint always need to be included in \(g\) in order to make sure
that \(p(n)\) sums to \(1\). The objective is then to find the values of
the Lagrange multipliers \(\lambda_i\) that optimize a function \(F\):

\begin{equation}\protect\hypertarget{eq:F}{}{F = H - \sum_{i=1}^m \lambda_i (g_i-c_i),}\label{eq:F}\end{equation}

where \(g_i\) is the mathematical formulation of the constraint \(i\)
and \(c_i\), its value. Note that \(F\) is just Shannon's entropy to
which we added terms that each sums to zero (\(g_i = c_i\)). \(F\) is
maximized by setting to \(0\) its partial derivative with respect to
\(p(n)\). In this contribution, we show how this can be done by deriving
the joint degree distribution analytically from the number of species
and the number of interactions in food webs.

\hypertarget{data-and-code}{%
\section{Data and code}\label{data-and-code}}

We tested our analytical and heuristic MaxEnt models against open
food-web data queried from three different sources and integrated into
what we call our \emph{complete dataset}. These sources include (1)
terrestrial and aquatic food webs sampled globally and archived on the
ecological interactions database Mangal; (2) freshwater stream food webs
from New Zealand; and (3) aquatic food webs from Tuesday Lake (Michigan,
United States). First, all food webs archived on \texttt{mangal.io}
(Poisot et al. 2016; Banville, Vissault, and Poisot 2021) were directly
queried from the database (\(n = 235\)). Most ecological networks
archived on Mangal are multilayer networks, i.e.~networks that describe
different types of interactions. We kept all networks whose interactions
were mainly of predation and herbivory types, and removed the largest
network (\(S = 714\)) for computational efficiency reasons. Then, to
this set we added food webs from two different sources: the New-Zealand
dataset (\(n = 21\); Pomeranz et al. 2018) and the Tuesday lake dataset
(\(n = 2\); Cohen, Jonsson, and Carpenter 2003). Of these two datasets,
\(19\) networks had data on species' relative abundances that were used
in the neutral model (Box 2). These \(19\) networks constitute our
\emph{abundance dataset}, which is a subset of our complete dataset.

All code and data to reproduce this article are available at the Open
Science Framework. Data cleaning, simulations and analyses were
conducted in Julia v1.5.4.

\hypertarget{analytical-models-measures-of-maximum-entropy}{%
\section{Analytical models: Measures of maximum
entropy}\label{analytical-models-measures-of-maximum-entropy}}

\hypertarget{joint-degree-distribution}{%
\subsection{Joint degree distribution}\label{joint-degree-distribution}}

The joint degree distribution \(p(k_{in},k_{out})\) is a joint discrete
probability distribution describing the probability that a species has
\(k_{in}\) predators and \(k_{out}\) preys, with \(k_{in}\) and
\(k_{out}\) \(\epsilon\) \([0, S]\). Basal species (e.g., plants) have a
\(k_{out}\) of \(0\), whereas top predators have a \(k_{in}\) of \(0\).
In contrast, the maximum number of preys and predators a species can
have is set by the number of species \(S\) in the food web. Here we show
how the joint degree distribution of maximum entropy can be obtained
given knowledge of \(S\) and \(L\).

We want to maximize Shannon's entropy

\begin{equation}\protect\hypertarget{eq:entropy_jdd}{}{H = -\sum_{k_{in}=0}^S\sum_{k_{out}=0}^S p(k_{in},k_{out}) \log p(k_{in},k_{out})}\label{eq:entropy_jdd}\end{equation}

subject to the following constraints:

\begin{equation}\protect\hypertarget{eq:g1}{}{g_1 = \sum_{k_{in}=0}^S\sum_{k_{out}=0}^S p(k_{in},k_{out}) = 1;}\label{eq:g1}\end{equation}

\begin{equation}\protect\hypertarget{eq:g2}{}{g_2 = \sum_{k_{in}=0}^S\sum_{k_{out}=0}^S k_{in} p(k_{in},k_{out}) = \langle k_{in} \rangle = \frac{L}{S};}\label{eq:g2}\end{equation}

\begin{equation}\protect\hypertarget{eq:g3}{}{g_3 = \sum_{k_{in}=0}^S\sum_{k_{out}=0}^S k_{out} p(k_{in},k_{out}) = \langle k_{out} \rangle = \frac{L}{S}.}\label{eq:g3}\end{equation}

The first constraint \(g_1\) is our normalizing constraint, whereas the
other two (\(g_2\) and \(g_3\)) fix the average of the marginal
distributions of \(k_{in}\) and \(k_{out}\) to the linkage density
\(L/S\). It is important to notice that
\(\langle k_{in} \rangle = \langle k_{out} \rangle\) because every edge
is associated to a predator and a prey. Therefore, without any further
constraints, we expect the joint degree distribution of maximum entropy
to be a symmetric probability distribution with regards to \(k_{in}\)
and \(k_{out}\). However, this does not mean that the joint degree
\emph{sequence} will be symmetric, since the joint degree sequence is
essentially a random realization of its probabilistic counterpart.

The joint probability distribution of maximum entropy given these
constraints is found using the method of Lagrange multipliers. To do so,
we seek to maximize the following expression:

\begin{equation}\protect\hypertarget{eq:F_jdd}{}{F = H - \lambda_1(g_1-1)-\lambda_2\left( g_2-\frac{L}{S}\right) - \lambda_3 \left( g_3-\frac{L}{S}\right),}\label{eq:F_jdd}\end{equation}

where \(\lambda_1\), \(\lambda_2\), and \(\lambda_3\) are the Lagrange
multipliers. The probability distribution that maximizes entropy is
obtained by finding these values. As pointed out in the introduction,
\(F\) is just Shannon's entropy to which we added terms that each sums
to zero (our constraints). \(F\) is maximized by setting to 0 its
partial derivative with respect to \(p(k_{in},k_{out})\). Because the
derivative of a constant is zero, this gives us:

\begin{equation}\protect\hypertarget{eq:lagrange_jdd}{}{\frac{\partial H}{\partial p(k_{in},k_{out})} = \lambda_1 \frac{\partial g_1}{\partial p(k_{in},k_{out})} + \lambda_2 \frac{\partial g_2}{\partial p(k_{in},k_{out})}+ \lambda_3 \frac{\partial g_3}{\partial p(k_{in},k_{out})}.}\label{eq:lagrange_jdd}\end{equation}

Evaluating the partial derivatives with respect to
\(p(k_{in},k_{out})\), we obtain:

\begin{equation}\protect\hypertarget{eq:lagrange2_jdd}{}{-\log p(k_{in},k_{out}) - 1 = \lambda_1 + \lambda_2 k_{in} + \lambda_3 k_{out}.}\label{eq:lagrange2_jdd}\end{equation}

Then, solving eq.~\ref{eq:lagrange2_jdd} for \(p(k_{in},k_{out})\), we
obtain:

\begin{equation}\protect\hypertarget{eq:lagrange3_jdd}{}{p(k_{in},k_{out}) = \frac{e^{-\lambda_2k_{in}-\lambda_3k_{out}}}{Z},}\label{eq:lagrange3_jdd}\end{equation}

where \(Z = e^{1+\lambda_1}\) is called the partition function. The
partition function ensures that probabilities sum to 1 (our
normalization constraint). It can be expressed in terms of \(\lambda_2\)
and \(\lambda_3\) as follows:

\begin{equation}\protect\hypertarget{eq:Z}{}{Z = \sum_{k_{in}=0}^S\sum_{k_{out}=0}^S e^{-\lambda_2k_{in}-\lambda_3k_{out}}.}\label{eq:Z}\end{equation}

After substituting \(p(k_{in},k_{out})\) in eq.~\ref{eq:g2} and
eq.~\ref{eq:g3}, we get a nonlinear system of two equations and two
unknowns:

\begin{equation}\protect\hypertarget{eq:lagrange4_jdd}{}{\frac{1}{Z}\sum_{k_{in}=0}^S\sum_{k_{out}=0}^S k_{in} e^{-\lambda_2k_{in}-\lambda_3k_{out}}  = \frac{L}{S};}\label{eq:lagrange4_jdd}\end{equation}

\begin{equation}\protect\hypertarget{eq:lagrange5_jdd}{}{\frac{1}{Z}\sum_{k_{in}=0}^S\sum_{k_{out}=0}^S k_{out} e^{-\lambda_2k_{in}-\lambda_3k_{out}}  = \frac{L}{S}.}\label{eq:lagrange5_jdd}\end{equation}

We solved eq.~\ref{eq:lagrange4_jdd} and eq.~\ref{eq:lagrange5_jdd}
numerically using the Julia library \texttt{JuMP.jl} v0.21.8 (Dunning,
Huchette, and Lubin 2017). \texttt{JuMP.jl} supports nonlinear
optimization problems by providing exact second derivatives that
increase the accuracy and performance of its solvers. The estimated
values of \(\lambda_2\) and \(\lambda_3\) can be substituted in
eq.~\ref{eq:lagrange3_jdd} to have a more workable expression for the
joint degree distribution.

We predicted the joint degree distribution of maximum entropy for each
food web in our complete dataset, i.e.~using their numbers of species
and numbers of interactions as state variables. We then sampled one
realization of the degree sequence for each network using the
probabilities given by the joint degree distribution. In
fig.~\ref{fig:joint_dd} (left panels), we show the relationship between
\(k_{out}\) and \(k_{in}\) standardized by the number of species in
their networks, for empirical and maximum entropy joint degree
distributions. We see that our model predicts higher values of
generality and vulnerability compared to empirical food webs
(i.e.~relative values of \(k_{out}\) and \(k_{in}\) closer to \(1\)).
However, plotting the difference between predicted and empirical values
for each species gives a different perspective. The right panel of
fig.~\ref{fig:joint_dd} presents these differences when species are
ordered by their total degree in their network (i.e.~by the sum of their
in and out-degrees). Indeed, our predicted joint degree sequences have
the same number of species as their empirical counterparts, but they are
species agnostic; in other words, instead of predicting a value for each
species directly, we predicted the entire joint degree sequence without
taking into account species' identity. When we associate predictions and
empirical data according to their rank in total degrees, we see that
species predicted to be have a higher generality (number of preys)
generally have a lower vulnerability (number of preys) than what is
observed (and conversely). In fig.~S1, we show how these differences
change when species are instead ordered by their out-degrees (left
panel) and in-degrees (right panel), respectively.

\begin{figure}
\hypertarget{fig:joint_dd}{%
\centering
\includegraphics{figures/joint_degree_dist.png}
\caption{Relative number of predators (\(k_{in}\)) as a function of the
relative number of preys (\(k_{out}\)) for each species in (a) empirical
and (b) maximum entropy joint degree sequences. Empirical networks
include all food webs archived on Mangal, as well as the New-Zealand and
Tuesday lake datasets (our complete dataset). The predicted joint degree
sequences were obtained after sampling one realization of the joint
degree distribution of maximum entropy for each network. (c) Difference
between predicted and empirical values when species are ordered
according to their total degrees. In all panels, each dot corresponds to
a single species in one of the network of our complete
dataset.}\label{fig:joint_dd}
}
\end{figure}

We plotted the Kullback--Leibler divergence between in and out-degree
distributions to compare the symmetry of empirical and maximum entropy
joint degree sequences fig.~\ref{fig:kl_diverg}. As we expected, our
model predicted more similar in-degree and out-degree distributions than
empirical data. However, this difference decreased with connectance
right panel of \ref{fig:kl_diverg}. Overall, this suggests that other
ecological constraints might be needed to account for the asymmetry of
the joint degree distribution, especially for networks with a lower
connectance.

\begin{figure}
\hypertarget{fig:kl_diverg}{%
\centering
\includegraphics{figures/kl_divergence.png}
\caption{(a) Probability density of KL divergence between in and
out-degree sequences of empirical and maximum entropy joint degree
sequences. (b) Difference between the KL divergence of empirical and
predicted networks as a function of connectance. In both panels,
empirical networks include all food webs archived on Mangal, as well as
the New-Zealand and Tuesday lake datasets (our complete dataset). The
joint degree sequence of simulated networks was obtained after sampling
one realization of the joint degree distribution of maximum entropy for
each network.}\label{fig:kl_diverg}
}
\end{figure}

\hypertarget{degree-distribution}{%
\subsection{Degree distribution}\label{degree-distribution}}

The degree distribution \(p(k)\) represents the probability that a
species has \(k\) interactions in a food web, with
\(k = k_{in} + k_{out}\). It can thus be directly obtained from the
joint degree distribution:

\[p(k) = \sum_{i=0}^k p(k_{in} = k - i, k_{out} = i).\]

In fig.~S2, we show that the degree distribution of maximum entropy,
given \(S\) and \(L\), predicts very low probabilities that a species
will be isolated in its food web (i.e.~having \(k = 0\)). As MacDonald,
Banville, and Poisot (2020) pointed out, the size of food webs should at
least be of \(S-1\) interactions, since a lower number would yield
isolated species, i.e.~species without any predators or preys. Our
results show that, under our purely information-theoretic model, the
probability that a species is isolated is quite high when the total
number of interactions is below this threshold. The expected proportion
of isolated species rapidly declines by orders of magnitude with
increasing numbers of species and interactions.

The degree distribution could also have been obtained directly using the
principle of maximum entropy, as discussed in Williams (2011). This
gives the following distribution:

\begin{equation}\protect\hypertarget{eq:lagrange_dd}{}{p(k) = \frac{e^{-\lambda_2k}}{Z},}\label{eq:lagrange_dd}\end{equation}

with \(Z = \sum_{k=0}^S e^{-\lambda_2k}.\)

This can be solved numerically using the constraint of the average
degree \(\langle k \rangle = \frac{2L}{S}\) of a species. Note that the
mean degree is twice the value of the linkage density, because every
link must be counted twice when we add in and out-degrees together.

\begin{equation}\protect\hypertarget{eq:lagrange2_dd}{}{\frac{1}{Z}\sum_{k=0}^S k e^{-\lambda_2k} = \frac{2L}{S}}\label{eq:lagrange2_dd}\end{equation}

The numerical solution is identical to the one we obtained using the
joint degree distribution as an intermediate. Ecologists wanting to
model a system without considering isolated species could simply change
the lower limit of \(k\) to \(1\) and solve the resulting equation
numerically.

In this section, we showed how important measures of food-web structure,
namely the degree distribution and the joint degree distribution, could
be derived with the principle of maximum entropy using minimal knowledge
on a biological community. This type of models, although useful to make
least-biased predictions on many network properties, can be hard to
apply for other measures. Indeed, there are dozens of measures of
network structure (Delmas et al. 2019) and many are not calculated with
mathematical equations, but with algorithms. Moreover, the applicability
of this method to empirical systems is limited by the state variables we
can actually measure and use. In the next section, we propose a more
flexible method to predict many measures of network structure
simultaneously, i.e.~by finding heuristically the configuration of the
network having maximum entropy given partial knowledge of its emerging
structure.

\hypertarget{box-1---working-with-predicted-numbers-of-interactions}{%
\section{Box 1 - Working with predicted numbers of
interactions}\label{box-1---working-with-predicted-numbers-of-interactions}}

Our models need information on the number of species and the number of
interactions. However, since the later is rarely estimated empirically,
ecologists might need to use predictive methods to estimate the total
number of interactions in a food web.

We used the flexible links model of MacDonald, Banville, and Poisot
(2020) to predict the number of interactions from the number of species.
The flexible links model, in contrast to other predictive models of the
number of interactions, incorporates meaningful ecological constraints
into the prediction of \(L\), namely the minimum \(S-1\) and maximum
\(S^2\) numbers of interactions in food webs. It estimates the
proportion of the \(S^2 - (S - 1)\) \emph{flexible links} that are
realized. More precisely, this model states that the number of
\emph{realized} flexible links (or interactions) \(L_{FL}\) in a food
web represents the number of realized interactions above the minimum
(i.e.~\(L = L_{FL} + S - 1\)) and is obtained from a beta-binomial
distribution with \(S^2 - (S - 1)\) trials and parameters
\(\alpha = \mu e^\phi\) and \(\beta = (1 - \mu) e^\phi\):

\begin{equation}\protect\hypertarget{eq:BB}{}{L_{FL} \sim \mathrm{BB}(S^2 - (S - 1), \mu e^\phi, (1 - \mu) e^\phi),}\label{eq:BB}\end{equation}

where \(\mu\) is the average probability across food webs that a
flexible link is realized, and \(\phi\) is the concentration parameter
around \(\mu\).

We fitted the flexible links model on all food webs in our complete
dataset, and estimated the parameters of eq.~\ref{eq:BB} using a
Hamiltonian Monte Carlo sampler with static trajectory (\(1\) chain and
\(3000\) iterations):

\begin{equation}\protect\hypertarget{eq:BBpost}{}{ [\mu, \phi| \textbf{L}, \textbf{S}] \propto \prod_{i = 1}^{m} \mathrm{BB}(L_i - (S_i - 1) | S_i^2 - (S_i - 1)), \mu e^{\phi}, (1 - \mu) e^\phi) \times \mathrm{B}(\mu| 3 , 7 ) \times \mathcal{N}(\phi | 3, 0.5), }\label{eq:BBpost}\end{equation}

where \(m\) is the number of food webs (\(m = 257\)) and \(\textbf{L}\)
and \(\textbf{S}\) are respectively the vectors of their numbers of
interactions and numbers of species. Our weakly-informative prior
distributions were chosen following MacDonald, Banville, and Poisot
(2020), i.e.~a beta distribution for \(\mu\) and a normal distribution
for \(\phi\). The Monte Carlo sampling of the posterior distribution was
conducted using the Julia library \texttt{Turing} v0.15.12.

The flexible links model is a generative model, i.e.~it can generate
plausible values of the predicted variable. We thus simulated \(1000\)
values of \(L\) for different values of \(S\) using the joint posterior
distribution of our model parameters (eq.~\ref{eq:BBpost}), and
calculated the mean degree for each simulated values. The resulting
distributions are shown in the left panel of
fig.~\ref{fig:degree_dist_fl} for three different values of species
richness. In the right panel of fig.~\ref{fig:degree_dist_fl}, we show
how the probability distribution for the mean degree constraints can be
used to generate a distribution of maximum entropy degree distributions,
since each simulated value of mean degree generates a different maximum
entropy degree distribution (eq.~\ref{eq:lagrange_dd} and
eq.~\ref{eq:lagrange2_dd}).

\begin{figure}
\hypertarget{fig:degree_dist_fl}{%
\centering
\includegraphics{figures/maxent_degree_dist_fl.png}
\caption{(a) Probability density of the predicted mean degree for
different values of species richness. The number of interactions was
predicted using the flexible links model fitted to all empirical
networks in our complete dataset (eq.~\ref{eq:BBpost}). (b) Degree
distributions of maximum entropy for a network of \(S=27\) species and
different numbers of interactions \(L\). The numbers of interactions
correspond to the lower and upper bounds of the \(67%
\), \(89%
\), and \(97%
\) percentile intervals (PI), as well as the median (in blue), of the
counterfactuals of the flexible links model. Each degree distribution
was obtained using eq.~\ref{eq:lagrange_dd} after solving numerically
eq.~\ref{eq:lagrange2_dd} using different values for the mean degree
constraint \(2L/S\).}\label{fig:degree_dist_fl}
}
\end{figure}

\hypertarget{box-2---corresponding-null-and-neutral-models}{%
\section{Box 2 - Corresponding null and neutral
models}\label{box-2---corresponding-null-and-neutral-models}}

\hypertarget{null-models-types-i-and-ii}{%
\subsection{Null models (types I and
II)}\label{null-models-types-i-and-ii}}

Our heuristic MaxEnt models, predicting adjacency matrices, were
compared with two topological null models. The first is the type I null
model of Fortuna and Bascompte (2006), in which the probability that a
species \(i\) predates on another species \(j\) is given by

\begin{equation}\protect\hypertarget{eq:type1null}{}{p(i \rightarrow j) = \frac{L}{S^2}.}\label{eq:type1null}\end{equation}

The second is the type II null model of Bascompte et al. (2003), in
which the probability of interaction is instead given by

\begin{equation}\protect\hypertarget{eq:type2null}{}{p(i \rightarrow j) = \frac{1}{2} \left(\frac{k_{in}(j)}{S} + \frac{k_{out}(i)}{S}\right),}\label{eq:type2null}\end{equation}

where \(k_{in}\) and \(k_{out}\) are the in and out-degrees,
respectively. The type I null model is based on connectance, whereas the
type II null model is based on the joint degree sequence. Therefore, the
type I and II topological null models correspond with our type I and II
MaxEnt network models, respectively, since they use similar constraints.

We predicted both types of null networks for all empirical networks in
our complete dataset (\(n = 257\)). We converted all probabilistic
networks to Boolean networks by generating \(100\) random Boolean
networks for each of these probabilistic webs. Then, we counted the
number of times each interaction was sampled, and kept the \(L\) entries
that were drawn the most amount of time, with \(L\) given by the number
of interactions in each food web. This ensured that the resulting null
networks had the same number of interactions as their empirical
counterparts.

\hypertarget{neutral-model}{%
\subsection{Neutral model}\label{neutral-model}}

We also compared our MaxEnt network models with a neutral model of
relative abundances, in which the probabilities of interaction are given
by

\begin{equation}\protect\hypertarget{eq:neutralmodel}{}{p(i \rightarrow j) \propto \frac{n_i}{N} \times \frac{n_j}{N},}\label{eq:neutralmodel}\end{equation}

where \(n_i\) and \(n_j\) are the abundances (or biomass) of both
species, and \(N\) is the total abundance (or biomass) of all species in
the network. We predicted neutral abundance matrices for all empirical
networks in our abundance dataset (\(n = 19\)), and converted these
weighted matrices to Boolean networks using an approach analogue to the
one we used for our null models.

\hypertarget{heuristic-models-networks-of-maximum-entropy}{%
\section{Heuristic models: Networks of maximum
entropy}\label{heuristic-models-networks-of-maximum-entropy}}

\hypertarget{from-shannons-to-svd-entropy}{%
\subsection{From Shannon's to SVD
entropy}\label{from-shannons-to-svd-entropy}}

The principle of maximum entropy can be applied on the network itself if
we decompose its adjacency matrix into a non-zero vector of relative
values. This is a necessary step when working with food webs, which are
frequently expressed as a matrix \(A = [a_{ij}]\) of Boolean values
representing the presence (\(a_{ij} = 1\)) or absence (\(a_{ij} = 0\))
of an interaction between two species \(i\) and \(j\). Knowing one or
many properties of a food web of interest (e.g., its number of species
and number of interactions), we can simulate its adjacency matrix
randomly by using these known ecological information to constrain the
space of potential networks. The entropy of this hypothetical matrix can
then be measured after decomposing it into appropriate values.
Simulating a series of networks until we find the one having the highest
entropy allows us to search for the most complex food-web configuration
given the ecological constraints used. This configuration is the least
biased one considering the information we had. In other words, the most
we can say about a network's adjacency matrix, without making more
assumptions than the ones given by our incomplete knowledge of its
structure, is the one of maximum entropy. Generating the most complex
network that corresponds to this structure allows us to explore more
easily other properties of food webs under MaxEnt.

Shannon's entropy can only be calculated on conventional probability
distributions such as the joint degree distribution. This is an issue
when working with ecological networks, which are represented as
adjacency matrices. For this reason, we need to use another measure of
entropy if we want to predict the network itself. We used the SVD
entropy as our measure of entropy, which is an application of Shannon's
entropy to the relative non-zero singular values of a truncated singular
value decomposition (t-SVD; Strydom, Dalla Riva, and Poisot 2021) of a
food web's Boolean adjacency matrix. We measured SVD entropy as follows:

\begin{equation}\protect\hypertarget{eq:svd-entropy}{}{J = -\sum_{i=1}^R s_i \log s_i,}\label{eq:svd-entropy}\end{equation}

where \(s_i\) are the relative singular values of the adjacency matrix
(\(s_i = \sigma_i / \sum_{i = 1}^R \sigma_i\), where \(\sigma_i\) are
the singular values). Note that the distribution of relative singular
values is analogous to a probability distribution, with \(0 < s_i < 1\)
and \(\sum s_i = 1\). Thus, this measure also satisfies all four
properties of an appropriate entropy measure (see introduction), while
being a proper measure of the internal complexity of food webs (Strydom,
Dalla Riva, and Poisot 2021). Following Strydom, Dalla Riva, and Poisot
(2021), we standardized this measure with the rank \(R\) of the matrix
(i.e.~\(J / \ln(R)\)) to account for the difference in dimensions
between networks (\emph{sensu} Pielou's evenness; Pielou 1975). We used
SVD entropy to predict the network configuration of maximum entropy
(i.e.~of maximum complexity) heuristically given a predetermined set of
constraints.

\hypertarget{maxent-network-models-types-i-and-ii}{%
\subsection{MaxEnt network models (types I and
II)}\label{maxent-network-models-types-i-and-ii}}

We built two types of MaxEnt network models: one based on connectance
(type I MaxEnt network model) and the other based on the joint degree
sequence (type II MaxEnt network model). They are based on the same
constraints as the types I and II null models presented above. For both
models, we used a simulated annealing algorithm with \(4\) chains,
\(2000\) steps and an initial temperature of \(0.2\). For each chain, we
first generated one random Boolean matrix with the same order (number of
species) as empirical webs, while maintaining the total number of
interactions (type I MaxEnt network model) or rows and columns sums
(type II MaxEnt network model). These are our initial configurations.
Then, we swapped interactions sequentially while maintaining the
original connectance or the joint degree sequence for types I and II
MaxEnt network model, respectively. Configurations with a higher SVD
entropy than the previous one in the chain were always accepted, whereas
they were accepted with a probability conditional to a decreasing
temperature when lower. The final configuration with the highest SVD
entropy among the four chains constitute our estimated MaxEnt network.
Even though we decided to work with point estimates, it is possible to
have a (non MaxEnt) probability distribution of networks when working
with the entire chains (or part of the chains). For each network in our
complete and abundance datasets, we estimated their configuration with
maximum entropy using both types of MaxEnt network models.

Our method is in contrast with maximum entropy graph models that predict
a probability distribution on networks under soft or hard constraints
(e.g., Park and Newman 2004; Cimini et al. 2019). We believe our
approach to be more flexible, easier to compute, while allowing direct
comparisons of empirical food webs with more complex networks with
similar structure.

\hypertarget{structure-of-maxent-networks}{%
\subsection{Structure of MaxEnt
networks}\label{structure-of-maxent-networks}}

We compared different measures of the structure of empirical food webs
with the ones of null, neutral, and MaxEnt networks
tbl.~\ref{tbl:measures_all}, and \ref{tbl:measures_abund}. For instance,
we measured the nestedness \(\rho\) according to Staniczenko, Kopp, and
Allesina (2013), i.e.~using the spectral radius of the adjacency matrix.
Nestedness indicates how much the diet of specialist species is a subset
of the one of generalists, and is strongly associated with network
modularity (Fortuna et al. 2010). We also measured the maximum trophic
level \emph{maxtl}, network diameter \emph{diam} (i.e.~the longest of
the shortest paths between all species pairs; Albert and Barabasi 2002),
the average maximum similarity between species pairs \emph{MxSim}, the
proportion of cannibal species \emph{Cannib} (i.e.~the proportion of
self loops), and the proportion of omnivorous species \emph{Omniv}
(i.e.~species whose preys are of different trophic levels).
\emph{MxSim}, \emph{Cannib}, and \emph{Omniv} are more deeply defined in
Williams and Martinez (2000).

Overall, we found that models based on the joint degree sequence (type
II null and MaxEnt network models) reproduced the structure of empirical
networks much better than the ones based on connectance and the neutral
model of relative abundances. On average, the type II MaxEnt network
model was better at predicting nestedness (\(0.62 \pm 0.08\)) than its
corresponding null model (\(0.73 \pm 0.05\); empirical networks:
\(0.63 \pm 0.09\)), as well as the proportion of cannibal species.
However, the type II null model was better at predicting network
diameter and average maximum similarity between species pairs.
Predictions were similar between both type II models for the maximum
trophic level and the proportion of omnivorous species.

\hypertarget{tbl:measures_all}{}
\begin{longtable}[]{@{}rrrrrrrr@{}}
\caption{\label{tbl:measures_all}Standardized mean differences of
predicted network measures with empirical data for all networks
(\(n = 257\)). Positive (negative) values indicate that the measure is
overestimated (underestimated) on average. Empirical networks are all
food webs archived on Mangal, as well as the New Zealand and Tuesday
lake food webs (our complete dataset). Null 1: Type I null model based
on connectance. MaxEnt 1: Type I MaxEnt network model based on
connectance. Null 2: Type II null model based on the joint degree
sequence. MaxEnt 2: Type II MaxEnt network model based on the joint
degree sequence. \(\rho\): nestedness measured by the spectral radius of
the adjacency matrix. \emph{maxtl}: maximum trophic level. \emph{diam}:
network diameter. \emph{MxSim}: average maximum similarity between
species pairs. \emph{Cannib}: proportion of cannibal species (self
loops). \emph{Omniv}: proportion of omnivorous species. \emph{entropy}:
SVD entropy.}\tabularnewline
\toprule
model & rho & maxtl & diam & MxSim & Cannib & Omniv &
entropy\tabularnewline
\midrule
\endfirsthead
\toprule
model & rho & maxtl & diam & MxSim & Cannib & Omniv &
entropy\tabularnewline
\midrule
\endhead
null 1 & -0.167 & 0.980 & 1.428 & -0.502 & 2.007 & 1.493 &
0.056\tabularnewline
MaxEnt 1 & -0.226 & 0.831 & 1.274 & -0.524 & 1.982 & 1.863 &
0.106\tabularnewline
null 2 & 0.160 & -0.125 & 0.016 & 0.007 & 1.078 & 0.559 &
-0.023\tabularnewline
MaxEnt 2 & -0.015 & 0.178 & 0.565 & -0.282 & 0.698 & 0.589 &
0.058\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{tbl:measures_abund}{}
\begin{longtable}[]{@{}rrrrrrrr@{}}
\caption{\label{tbl:measures_abund}Standardized mean differences of
predicted network measures with empirical data for networks having
abundance data (\(n = 19\)). Positive (negative) values indicate that
the measure is overestimated (underestimated) on average. Empirical
networks are all New Zealand and Tuesday lake food webs with abundance
data (our abundance dataset). Null 1: Type I null model based on
connectance. MaxEnt 1: Type I MaxEnt network model based on connectance.
Null 2: Type II null model based on the joint degree sequence. MaxEnt 2:
Type II MaxEnt network model based on the joint degree sequence.
\(\rho\): nestedness measured by the spectral radius of the adjacency
matrix. \emph{maxtl}: maximum trophic level. \emph{diam}: network
diameter. \emph{MxSim}: average maximum similarity between species
pairs. \emph{Cannib}: proportion of cannibal species (self loops).
\emph{Omniv}: proportion of omnivorous species. \emph{entropy}: SVD
entropy.}\tabularnewline
\toprule
model & rho & maxtl & diam & MxSim & Cannib & Omniv &
entropy\tabularnewline
\midrule
\endfirsthead
\toprule
model & rho & maxtl & diam & MxSim & Cannib & Omniv &
entropy\tabularnewline
\midrule
\endhead
neutral & 0.367 & -0.090 & 0.027 & 0.266 & 6.870 & 0.576 &
-0.083\tabularnewline
null 1 & -0.134 & 0.950 & 1.919 & -0.369 & 2.077 & 0.614 &
0.068\tabularnewline
MaxEnt 1 & -0.229 & 1.020 & 1.946 & -0.355 & 2.215 & 0.801 &
0.121\tabularnewline
null 2 & 0.128 & -0.115 & -0.135 & 0.157 & 1.444 & 0.029 &
-0.021\tabularnewline
MaxEnt 2 & -0.010 & 0.054 & 0.243 & -0.062 & -0.038 & 0.083 &
0.038\tabularnewline
\bottomrule
\end{longtable}

The picture slightly changes when we consider another important property
of ecological networks, i.e.~their motifs profile. We measured the
proportion of three-species motifs, which can be considered as simple
building blocks of ecological networks (Milo et al. 2002; Stouffer et
al. 2007). Motifs are the backbone of complex ecological networks from
which network structure is built upon and play a crucial role in
community dynamics and assembly (Bastolla et al. 2009; Stouffer and
Bascompte 2011). In fig.~\ref{fig:motifs}, we show that the motifs
profile of networks generated using the type II MaxEnt network model was
very close to the one of empirical data, and that they made better
predictions than the type II null model. This is also shown in
fig.~\ref{fig:motifs_rel}, where we see that relationships between
motifs pairs of empirical food webs are very similar to the ones of the
type II MaxEnt networks. Overall, these results suggest that the type II
MaxEnt network model can reproduce many aspects of network structure.
This highlights the importance of the joint degree sequence in shaping
ecological networks.

\begin{figure}
\hypertarget{fig:motifs}{%
\centering
\includegraphics{figures/motifs_distribution.png}
\caption{Motifs profile of empirical, maximum entropy, and null food
webs. Empirical networks include all food webs archived on Mangal, as
well as the New-Zealand and Tuesday lake datasets (our complete
dataset). Maximum entropy networks were estimated using a simulated
annealing algorithm to find the network of maximum SVD entropy while
maintaining the connectance (type I MaxEnt network model) or the joint
degree sequence (type II MaxEnt network model). The predictions of the
types I and II null models are also plotted. Boxplots display the median
proportions of each motif (middle horizontal lines), as well as the
first (bottom horizontal lines) and third (top horizontal lines)
quartiles. Vertical lines encompass all data points that fall within
\(1.5\) times the interquartile range from both quartiles, and dots are
data points that fall outside this range. Motifs names are from Stouffer
et al. (2007).}\label{fig:motifs}
}
\end{figure}

\begin{figure}
\hypertarget{fig:motifs_rel}{%
\centering
\includegraphics{figures/motifs_relations.png}
\caption{Relationship between motifs proportions of empirical, maximum
entropy, and null food webs. Empirical networks include all food webs
archived on Mangal, as well as the New-Zealand and Tuesday lake datasets
(our complete dataset). Maximum entropy networks were estimated using a
simulated annealing algorithm to find the network of maximum SVD entropy
while maintaining the connectance (type I MaxEnt network model) or the
joint degree sequence (type II MaxEnt network model). The predictions of
the types I and II null models are also plotted. Regression lines are
plotted in each panel. Motifs names are from Stouffer et al.
(2007).}\label{fig:motifs_rel}
}
\end{figure}

Moreover, we found that empirical networks are close to their maximum
entropy considering a fixed joint degree sequence (fig.~S3). Empirical
networks had an SVD entropy of \(0.89 \pm 0.04\), compared to type II
MaxEnt networks which had values of \(0.94 \pm 0.03\). The relationship
between the SVD entropy of empirical food webs and their maximum entropy
configuration is plotted in the last panel of fig.~\ref{fig:measures}.
We also found no relationship between the difference in SVD entropy
between empirical and type II MaxEnt networks and species richness, the
number of interactions, or connectance (fig.~S4). Similarly, we found no
correlation between the difference in SVD entropy and the Jaccard
distance of empirical and type II MaxEnt networks.

Finally, in fig.~\ref{fig:measures} we also show how well empirical
measures are predicted using the type II MaxEnt network model. In
accordance with our previous results, we found that nestedness was very
well predicted by our model. However, the type II MaxEnt network model
overestimated the maximum trophic level and network diameter, especially
for networks with high empirical values. In fig.~S5, we show that the
pairwise relationships between these four measures and species richness
in empirical food webs are similar (in magnitude and sign) to the ones
in type II MaxEnt networks.

\begin{figure}
\hypertarget{fig:measures}{%
\centering
\includegraphics{figures/measures_emp_maxent.png}
\caption{Relationship between the structure of empirical and maximum
entropy food webs. Empirical networks include all food webs archived on
Mangal, as well as the New-Zealand and Tuesday lake datasets (our
complete dataset). Maximum entropy networks were estimated using a
simulated annealing algorithm to find the network of maximum SVD entropy
while maintaining the joint degree sequence (type II MaxEnt network
model). (a) Nestedness (estimated using the spectral radius of the
adjacency matrix), (b) the maximum trophic level, (c) the network
diameter, and (d) the SVD entropy were measured on these empirical and
maximum entropy food webs. The identity line is plotted in each
panel.}\label{fig:measures}
}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The principle of maximum entropy is a robust mathematical method of
finding least-biased probability distributions that have some desired
properties given by prior knowledge on a system. We first used MaxEnt to
predict the joint degree distribution and the degree distribution of
maximum entropy given known numbers of species and of interactions. We
found that the resulting joint degree distributions were more symmetric
than the ones of empirical food webs, which suggests that other
constraints might be needed to improve those predictions. We also used
MaxEnt to predict networks of maximum entropy with a specified
structure. These networks are the most complex, or random, given the
constraints used. Likewise, we found that knowledge of species richness
and of the number of interactions were not sufficient to reproduce many
aspects of network structure found in nature. However, a model based on
the entire joint degree sequence, i.e.~on the number of preys and
predators for each species, gave more convincing results. Our type II
MaxEnt network model yielded better or similar predictions than the type
II null model, also based on the joint degree sequence, for most
measures considered including the motifs profile. This suggests that the
joint degree sequence drives many aspects of network structure. Indeed,
considering our findings that empirical networks are close to maximum
entropy for a given joint degree sequence, our results suggest that
food-web topology is configured almost entirely randomly around these
marginal numbers of predators and preys.

Our method and results could be used for different purposes. First, they
could be used as first-order approximations of network structure when
only state variables are known. This could prove useful when predicting
network structure at large spatial scales, where few ecological
information is known at that scale. Second, they could be used as
informative priors in Bayesian analyses of the structure of ecological
networks (e.g., Cirtwill et al. 2019). Third, they could be used to make
better predictions of pairwise species interactions by constraining the
space of feasible networks, as discussed in Strydom et al. (2021).
Finally, they could be used as alternative null models of ecological
networks to better understand ecological mechanisms driving food-web
structure. In that case, our model might need to be slightly adapted to
give a probability distribution of Boolean networks (in contrast with
point estimates of maximum entropy networks).

One of the biggest challenges in using the principle of maximum entropy
is to identify the set of state variables that best reproduce empirical
data. We found that the numbers of preys and predators for each species
are important state variables for the prediction of maximum entropy
networks. However, our predictions overestimated some measures of
network structure, especially the maximum trophic level and network
diameter. Therefore, we should continue playing the ecological detective
to find these other topological constraints that would improve the
predictions of our MaxEnt network models.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. This
work was supported by the Institute for Data Valorisation (IVADO) and
the Natural Sciences and Engineering Research Council of Canada (NSERC)
Collaborative Research and Training Experience (CREATE) program, through
the Computational Biodiversity Science and Services (BIOS\(^2\))
program.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Albert2002StaMec}{}%
Albert, R., and A. L. Barabasi. 2002. {``Statistical Mechanics of
Complex Networks.''} \emph{Reviews of Modern Physics} 74 (1): 47--97.
\url{https://doi.org/10.1103/RevModPhys.74.47}.

\leavevmode\hypertarget{ref-Banville2021ManJla}{}%
Banville, Francis, Steve Vissault, and Timothée Poisot. 2021.
{``Mangal.jl and EcologicalNetworks.jl: Two Complementary Packages for
Analyzing Ecological Networks in Julia.''} \emph{Journal of Open Source
Software} 6 (61): 2721. \url{https://doi.org/10.21105/joss.02721}.

\leavevmode\hypertarget{ref-Bascompte2003NesAssa}{}%
Bascompte, J., P. Jordano, C. J. Melian, and J. M. Olesen. 2003. {``The
Nested Assembly of Plant-Animal Mutualistic Networks.''}
\emph{Proceedings of the National Academy of Sciences of the United
States of America} 100 (16): 9383--87.
\url{https://doi.org/10.1073/pnas.1633576100}.

\leavevmode\hypertarget{ref-Bastolla2009ArcMut}{}%
Bastolla, Ugo, Miguel A. Fortuna, Alberto Pascual-Garcia, Antonio
Ferrera, Bartolo Luque, and Jordi Bascompte. 2009. {``The Architecture
of Mutualistic Networks Minimizes Competition and Increases
Biodiversity.''} \emph{Nature} 458 (7241): 1018--U91.
\url{https://doi.org/10.1038/nature07950}.

\leavevmode\hypertarget{ref-Beck2009GenInf}{}%
Beck, Christian. 2009. {``Generalised Information and Entropy Measures
in Physics.''} \emph{Contemporary Physics} 50 (4): 495--510.
\url{https://doi.org/10.1080/00107510902823517}.

\leavevmode\hypertarget{ref-Cimini2019StaPhy}{}%
Cimini, Giulio, Tiziano Squartini, Fabio Saracco, Diego Garlaschelli,
Andrea Gabrielli, and Guido Caldarelli. 2019. {``The Statistical Physics
of Real-World Networks.''} \emph{Nature Reviews Physics} 1 (1): 58--71.
\url{https://doi.org/10.1038/s42254-018-0002-6}.

\leavevmode\hypertarget{ref-Cirtwill2019QuaFrac}{}%
Cirtwill, Alyssa R., Anna Eklof, Tomas Roslin, Kate Wootton, and
Dominique Gravel. 2019. {``A Quantitative Framework for Investigating
the Reliability of Empirical Network Construction.''} \emph{Methods in
Ecology and Evolution} 10 (6): 902--11.
\url{https://doi.org/10.1111/2041-210X.13180}.

\leavevmode\hypertarget{ref-Cohen2003EcoComa}{}%
Cohen, Joel E., Tomas Jonsson, and Stephen R. Carpenter. 2003.
{``Ecological Community Description Using the Food Web, Species
Abundance, and Body Size.''} \emph{Proceedings of the National Academy
of Sciences} 100 (4): 1781--86.
\url{https://doi.org/10.1073/pnas.232715699}.

\leavevmode\hypertarget{ref-Delmas2019AnaEco}{}%
Delmas, Eva, Mathilde Besson, Marie Hélène Brice, Laura A. Burkle,
Giulio V. Dalla Riva, Marie Josée Fortin, Dominique Gravel, et al. 2019.
{``Analysing Ecological Networks of Species Interactions.''}
\emph{Biological Reviews}. \url{https://doi.org/10.1111/brv.12433}.

\leavevmode\hypertarget{ref-Dunning2017JumMod}{}%
Dunning, Iain, Joey Huchette, and Miles Lubin. 2017. {``JuMP: A Modeling
Language for Mathematical Optimization.''} \emph{SIAM Review} 59 (2):
295--320. \url{https://doi.org/10.1137/15M1020575}.

\leavevmode\hypertarget{ref-Fortuna2006HabLos}{}%
Fortuna, M. A., and J. Bascompte. 2006. {``Habitat Loss and the
Structure of Plant-Animal Mutualistic Networks.''} \emph{Ecology
Letters} 9 (3): 278--83.
\url{https://doi.org/10.1111/j.1461-0248.2005.00868.x}.

\leavevmode\hypertarget{ref-Fortuna2010NesMod}{}%
Fortuna, M. A., Daniel B. Stouffer, Jens M. Olesen, Pedro Jordano, David
Mouillot, Boris R. Krasnov, Robert Poulin, and Jordi Bascompte. 2010.
{``Nestedness Versus Modularity in Ecological Networks: Two Sides of the
Same Coin?''} \emph{Journal of Animal Ecology} 79 (4): 811--17.
\url{https://doi.org/10.1111/j.1365-2656.2010.01688.x}.

\leavevmode\hypertarget{ref-Frank2011SimDera}{}%
Frank, S. A., and E. Smith. 2011. {``A Simple Derivation and
Classification of Common Probability Distributions Based on Information
Symmetry and Measurement Scale.''} \emph{Journal of Evolutionary
Biology} 24 (3): 469--84.
\url{https://doi.org/10.1111/j.1420-9101.2010.02204.x}.

\leavevmode\hypertarget{ref-Harremoes2001MaxEnt}{}%
Harremoës, Peter, and Flemming Topsøe. 2001. {``Maximum Entropy
Fundamentals.''} \emph{Entropy} 3 (3): 191--226.
\url{https://doi.org/10.3390/e3030191}.

\leavevmode\hypertarget{ref-Harte2014MaxInf}{}%
Harte, J., and Erica A. Newman. 2014. {``Maximum Information Entropy: A
Foundation for Ecological Theory.''} \emph{Trends in Ecology \&
Evolution} 29 (7): 384--89.
\url{https://doi.org/10.1016/j.tree.2014.04.009}.

\leavevmode\hypertarget{ref-Harte2008MaxEnt}{}%
Harte, J., T. Zillio, E. Conlisk, and A. B. Smith. 2008. {``Maximum
Entropy and the State-Variable Approach to Macroecology.''}
\emph{Ecology} 89 (10): 2700--2711.
\url{https://doi.org/10.1890/07-1369.1}.

\leavevmode\hypertarget{ref-vanderHoorn2018SpaMaxa}{}%
Hoorn, Pim van der, Gabor Lippner, and Dmitri Krioukov. 2018. {``Sparse
Maximum-Entropy Random Graphs with a Given Power-Law Degree
Distribution.''} \emph{Journal of Statistical Physics} 173 (3): 806--44.
\url{https://doi.org/10.1007/s10955-017-1887-7}.

\leavevmode\hypertarget{ref-Hortal2015SevSho}{}%
Hortal, Joaquín, Francesco de Bello, José Alexandre F. Diniz-Filho,
Thomas M. Lewinsohn, Jorge M. Lobo, and Richard J. Ladle. 2015. {``Seven
Shortfalls That Beset Large-Scale Knowledge of Biodiversity.''}
\emph{Annual Review of Ecology, Evolution, and Systematics} 46 (1):
523--49. \url{https://doi.org/10.1146/annurev-ecolsys-112414-054400}.

\leavevmode\hypertarget{ref-Jaynes1957InfThe}{}%
Jaynes, E. T. 1957a. {``Information Theory and Statistical Mechanics.''}
\emph{Physical Review} 106 (4): 620--30.
\url{https://doi.org/10.1103/PhysRev.106.620}.

\leavevmode\hypertarget{ref-Jaynes1957InfThea}{}%
---------. 1957b. {``Information Theory and Statistical Mechanics.
II.''} \emph{Physical Review} 108 (2): 171--90.
\url{https://doi.org/10.1103/PhysRev.108.171}.

\leavevmode\hypertarget{ref-Khinchin2013MatFou}{}%
Khinchin, A. Ya. 2013. \emph{Mathematical Foundations of Information
Theory}. Courier Corporation.

\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
MacDonald, Arthur Andrew Meahan, Francis Banville, and Timothée Poisot.
2020. {``Revisiting the Links-Species Scaling Relationship in Food
Webs.''} \emph{Patterns} 0 (0).
\url{https://doi.org/10.1016/j.patter.2020.100079}.

\leavevmode\hypertarget{ref-Martyushev2006MaxEnt}{}%
Martyushev, L. M., and V. D. Seleznev. 2006. {``Maximum Entropy
Production Principle in Physics, Chemistry and Biology.''} \emph{Physics
Reports-Review Section of Physics Letters} 426 (1): 1--45.
\url{https://doi.org/10.1016/j.physrep.2005.12.001}.

\leavevmode\hypertarget{ref-Milo2002NetMot}{}%
Milo, R., S. Shen-Orr, S. Itzkovitz, N. Kashtan, D. Chklovskii, and U.
Alon. 2002. {``Network Motifs: Simple Building Blocks of Complex
Networks.''} \emph{Science} 298 (5594): 824--27.
\url{https://doi.org/10.1126/science.298.5594.824}.

\leavevmode\hypertarget{ref-Park2004StaMeca}{}%
Park, Juyong, and M. E. J. Newman. 2004. {``Statistical Mechanics of
Networks.''} \emph{Physical Review E} 70 (6): 066117.
\url{https://doi.org/10.1103/PhysRevE.70.066117}.

\leavevmode\hypertarget{ref-Phillips2006MaxEnta}{}%
Phillips, Steven J., Robert P. Anderson, and Robert E. Schapire. 2006.
{``Maximum Entropy Modeling of Species Geographic Distributions.''}
\emph{Ecological Modelling} 190 (3): 231--59.
\url{https://doi.org/10.1016/j.ecolmodel.2005.03.026}.

\leavevmode\hypertarget{ref-Pielou1975EcoDiv}{}%
Pielou, Evelyn C. 1975. {``Ecological Diversity.''} In. 574.524018 P5.

\leavevmode\hypertarget{ref-Poisot2016ManMakb}{}%
Poisot, Timothee, Benjamin Baiser, Jennifer A. Dunne, Sonia Kefi,
Francois Massol, Nicolas Mouquet, Tamara N. Romanuk, Daniel B. Stouffer,
Spencer A. Wood, and Dominique Gravel. 2016. {``Mangal - Making
Ecological Network Analysis Simple.''} \emph{Ecography} 39 (4): 384--90.
\url{https://doi.org/10.1111/ecog.00976}.

\leavevmode\hypertarget{ref-Pomeranz2018DatInf}{}%
Pomeranz, Justin Page, Ross M. Thompson, Timothée Poisot, Jon S.
Harding, and Justin P. F. Pomeranz. 2018. {``Data from: Inferring
Predator-Prey Interactions in Food Webs.''} Dryad.
\url{https://doi.org/10.5061/DRYAD.K59M37F}.

\leavevmode\hypertarget{ref-Shannon1948MatThe}{}%
Shannon, C. E. 1948. {``A Mathematical Theory of Communication.''}
\emph{The Bell System Technical Journal} 27 (3): 379--423.
\url{https://doi.org/10.1002/j.1538-7305.1948.tb01338.x}.

\leavevmode\hypertarget{ref-Staniczenko2013GhoNes}{}%
Staniczenko, Phillip P. A., Jason C. Kopp, and Stefano Allesina. 2013.
{``The Ghost of Nestedness in Ecological Networks.''} \emph{Nature
Communications} 4 (1): 1391. \url{https://doi.org/10.1038/ncomms2422}.

\leavevmode\hypertarget{ref-Stock2021OptTra}{}%
Stock, Michiel, Timothée Poisot, and Bernard De Baets. 2021. {``Optimal
Transportation Theory for Species Interaction Networks.''} \emph{Ecology
and Evolution} 11 (9): 3841--55.
\url{https://doi.org/10.1002/ece3.7254}.

\leavevmode\hypertarget{ref-Stouffer2011ComInc}{}%
Stouffer, Daniel B., and Jordi Bascompte. 2011. {``Compartmentalization
Increases Food-Web Persistence.''} \emph{Proceedings of the National
Academy of Sciences} 108 (9): 3648--52.
\url{https://doi.org/10.1073/pnas.1014353108}.

\leavevmode\hypertarget{ref-Stouffer2007EviExi}{}%
Stouffer, Daniel B., Juan Camacho, Wenxin Jiang, and Luis A. Nunes
Amaral. 2007. {``Evidence for the Existence of a Robust Pattern of Prey
Selection in Food Webs.''} \emph{Proceedings of the Royal Society
B-Biological Sciences} 274 (1621): 1931--40.
\url{https://doi.org/10.1098/rspb.2007.0571}.

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, Tanya, Michael D. Catchen, Francis Banville, Dominique Caron,
Gabriel Dansereau, Philippe Desjardins-Proulx, Norma R. Forero-Muñoz, et
al. 2021. {``A Roadmap Towards Predicting Species Interaction Networks
(across Space and Time).''} \emph{Philosophical Transactions of the
Royal Society B: Biological Sciences} 376 (1837): 20210063.
\url{https://doi.org/10.1098/rstb.2021.0063}.

\leavevmode\hypertarget{ref-Strydom2021SvdEnt}{}%
Strydom, Tanya, Giulio V. Dalla Riva, and Timothee Poisot. 2021. {``SVD
Entropy Reveals the High Complexity of Ecological Networks.''}
\emph{Frontiers in Ecology and Evolution} 9: 623141.
\url{https://doi.org/10.3389/fevo.2021.623141}.

\leavevmode\hypertarget{ref-Williams2011BioMet}{}%
Williams, Richard J. 2011. {``Biology, Methodology or Chance? The Degree
Distributions of Bipartite Ecological Networks.''} \emph{PLOS ONE} 6
(3): e17645. \url{https://doi.org/10.1371/journal.pone.0017645}.

\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
Williams, Richard J., and Neo D. Martinez. 2000. {``Simple Rules Yield
Complex Food Webs.''} \emph{Nature} 404 (6774): 180--83.
\url{https://doi.org/10.1038/35004572}.

\end{CSLReferences}

\end{document}